%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Master Thesis - HPC Sorting Serious Game
% University of Genova
%
% Author: Szymon Zinkowicz
% Supervisor: Professor Daniele D'Agostino
% Reviewer: Professor Maura Cerioli
%
% License: MIT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[twoside]{./class/masterthesis}

%-----------------------------------------------------------------------
%	PACKAGES
%-----------------------------------------------------------------------

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[table, dvipsnames]{xcolor}
\usepackage[backend=biber,style=ieee,sorting=none,natbib=true]{biblatex}
\usepackage{hyperref}
\usepackage{subcaption}
\usepackage{caption}
\usepackage{lipsum}
\usepackage{fancyhdr}
\usepackage{float}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{array,multirow,makecell}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{tocloft}
\usepackage{mdframed}
\usepackage{microtype}

%-----------------------------------------------------------------------
%	GEOMETRY
%-----------------------------------------------------------------------

\geometry{
	top=3cm,
	bottom=3cm,
	left=3.5cm,
	right=2.5cm,
	headheight=15pt
}

%-----------------------------------------------------------------------
%	HYPERREF SETUP
%-----------------------------------------------------------------------

\hypersetup{
	colorlinks=true,
	linktoc=all,
	linkcolor=blue,
	citecolor=blue,
	urlcolor=blue,
	pdfauthor={Szymon Zinkowicz},
	pdftitle={HPC Sorting Serious Game: A Web-First Educational Approach},
}

%-----------------------------------------------------------------------
%	CODE LISTING SETUP
%-----------------------------------------------------------------------
  
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
 
\lstdefinestyle{gdscript}{
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    belowcaptionskip=1\baselineskip,
    frame=none,
    basicstyle=\footnotesize\ttfamily,
    keywordstyle=\bfseries\color{green!40!black},
    commentstyle=\itshape\color{purple!40!black},
    identifierstyle=\color{blue},
    backgroundcolor=\color{gray!10!white},
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=4,
    language=Python, % GDScript is Python-like
    mathescape=false
}

\lstset{style=gdscript}

%-----------------------------------------------------------------------
%	BIBLIOGRAPHY
%-----------------------------------------------------------------------

\addbibresource{references.bib}

%-----------------------------------------------------------------------
%	CUSTOM MACROS
%-----------------------------------------------------------------------

% Convenient macros for frequently used terms
\newcommand*\openmp{OpenMP}
\newcommand*\mpi{MPI}
\newcommand*\gdscript{GDScript}
\newcommand*\hpc{HPC}

%-----------------------------------------------------------------------
% Measurement placeholders — replace \MEASURE{placeholder} with actual values
% after running benchmarks. Search for \MEASURE to find all unfilled values.
\newcommand{\MEASURE}[1]{\textbf{[\,#1\,]}}  % renders as bold [placeholder]

% --- Web Export Size (measured via Get-ChildItem + Python gzip) ---
\newcommand{\exportSizeUncompressed}{135}           % MB, uncompressed export folder (actual: 135.32 MB)
\newcommand{\exportSizeGzip}{30}                    % MB, gzip-compressed (actual: 30.38 MB at level 6)
\newcommand{\exportSizeWasm}{40}                    % MB, index.side.wasm=38.88 + index.wasm=1.52 = 40.4 MB
\newcommand{\exportSizeScenes}{91}                  % MB, index.pck=90.4 MB (bundles scenes, scripts, assets)
\newcommand{\exportSizeAssets}{\MEASURE{??}}       % MB, subset of .pck — not separately measurable without Godot PCK tool
\newcommand{\exportSizeGDSync}{\MEASURE{??}}       % MB, subset of .pck — not separately measurable without Godot PCK tool


%-----------------------------------------------------------------------
%	BEGIN DOCUMENT
%-----------------------------------------------------------------------

\begin{document}

% Document information for masterthesis.cls
\title{A serious game for High Performance Computing}
\author{Szymon Zinkowicz}
\advisor{Prof. Daniele D'Agostino}
% \examiner{Prof. Maura Cerioli}
\examiner{Prof. Gianna Reggio}

% Generate title page using masterthesis.cls
\maketitle

%-----------------------------------------------------------------------
%	DEDICATION (Optional)
%-----------------------------------------------------------------------

\dedication{
    \textit{To my father Piotr,\\
        from whom I learned modesty, hard work, and integrity, and whose sacrifices made this possible.}

    \vspace{1.5\baselineskip}

    \textit{To my mother Wioletta,\\
        who always believed in me and pushed me forward.}

    \vspace{1.5\baselineskip}

    %     \textit{To my girlfriend Magdalena,\\
    %         who has always been focused and keeping me on my toes.}
}

%-----------------------------------------------------------------------
%	ABSTRACT
%-----------------------------------------------------------------------

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

High Performance Computing (HPC) and parallel programming pose significant pedagogical challenges, as students must internalize abstract concepts without tangible interactive experiences. Serious games facilitate intuitive understanding through engagement, surfacing misconceptions and enabling targeted instruction.

This thesis presents a web-based educational serious game for teaching fundamental parallel computing concepts—specifically OpenMP shared memory parallelism—through interactive card-sorting mechanics. Players sort cards collaboratively in multiplayer mode using local buffer zones, simulating OpenMP thread-local storage. Students cooperate to minimize sorting time under constraints, learning speedup, scalability, communication overhead, and multithreading.

Key contributions include: (1) an original mapping between card-sorting mechanics and HPC concepts derived from physical classroom experiments, (2) robust multiplayer state synchronization using GDSync, (3) responsive UI/UX for manipulating 50+ interactive elements across screen sizes, (4) documentation of technical challenges in GDSync and multiplayer synchronization, and (5) an open-source, extensible platform for HPC education research.

\textbf{Keywords:} Serious Games, High Performance Computing, Parallel Computing Education, OpenMP, Web-Based Game Development, Multiplayer Synchronization, Godot Engine, GDScript, Card Sorting, Educational Technology

%-----------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%-----------------------------------------------------------------------

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to express my sincere gratitude to my supervisor, Professor Daniele D'Agostino, for his guidance, valuable insights, and continuous support throughout this thesis work. His much needed enthusiasm was what has put me on the rails for High Performance Computing research. His expertise in High Performance Computing and educational methodologies has been instrumental in shaping this research.

I am also grateful to Professor Maura Cerioli for serving as the reviewer of this thesis for the first period, steering me in right direction regarding organisational aspects and providing constructive feedback.

Special thanks to the Godot Engine community and the developers of the GDSync framework for their open-source contributions and support in resolving technical issues encountered during development.

Finally, I thank my family and friends for their encouragement and patience during the completion of this work.

%-----------------------------------------------------------------------
%	TABLE OF CONTENTS
%-----------------------------------------------------------------------

\tableofcontents
\listoffigures
\listoftables
\lstlistoflistings


%	MAIN CONTENT
%-----------------------------------------------------------------------

\input{chapters/01-introduction}
\input{chapters/02-background}
\input{chapters/03-methodology}
\input{chapters/04-architecture}
\input{chapters/05-implementation}
\input{chapters/06-results}
\input{chapters/07-problems}
\input{chapters/08-conclusion}


%	BIBLIOGRAPHY
%-----------------------------------------------------------------------

\printbibliography[heading=bibintoc]


%	APPENDICES
%-----------------------------------------------------------------------

\appendix

\chapter{User Manual}
\input{appendices/appendix-a-user-manual}

\chapter{Code Listings}
\input{appendices/appendix-b-code-listings}

\chapter{API Documentation}
\input{appendices/appendix-c-api-documentation}

%TODO - add study materials (slides, exercises, etc.) as appendix D after the review
% \chapter{Study Materials}
% \input{appendices/appendix-d-study-materials}

%-----------------------------------------------------------------------

\end{document}
